---
title: "STAT3550_Lab1"
author: "Luke Rodriguez"
date: "2026-01-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(ISLR2)
```
Fit a simple linear regressio nmodel using the botton dataset, with median house value as the response variable low economic status as the explanatory variable.
```{r}
head(Boston)
attach(Boston)
lm.fit <- lm(medv ~ lstat) # y ~ x
```

view the output of the model using the summary function
```{r}
lm.fit
summary(lm.fit)

```

Coefficient
```{r}
coef(lm.fit)
```

Confidence intervals 
Use the confint function to create a 95% CI
```{r}
confint(lm.fit)
```

Do predictions and create prediction intervals based on 3 new values of lstat variable (5, 10, and 15).
```{r}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "prediction")
```

visualizing the model
```{r}
plot(lstat, medv)
abline(lm.fit)
```

Notice that the model seems relatively linear, but may show some nonlinearity, particularly seen in lower values values of lstat.

checking model assumptions
```{r}
par(mfrow = c(2,2))
plot(lm.fit)
```

Notice that the residuals are not exhibiting constant variance, which might indicate that the model assumptions are not satisfied. Perhaps a different model will work etter for our data.

Multiple linear gression

Fit a model with two predictors
```{r}
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

Fit a model with all predictors
```{r}
lm.fit <- lm(medv ~., data = Boston) # take all the other variables as the explanatory variables (if leave x out)
summary(lm.fit)
```

nonlinear transformations

using the boston housing dataset, investigate whether the relationship between median house value(medv) and percentage 

```{r}
```

polynomial moel

fit a quadratic regression model
```{r}
#linear model 
lm_linear <- lm(medv ~ lstat, data = Boston)
summary(lm_linear)

#quadratic model
lm_quad <- lm(medv ~ lstat + I(lstat^2), data = Boston)
summary(lm_quad)
```

Run anova test to see if the model improves form the original linear model when adding the quadratic term.
```{r}
anova(lm_linear, lm_quad)
```

```{r}
par(mfrow = c(2,2)) # use par to compare parallel
plot(lm_quad)
```

log linear model
```{r}
#scatterplot
```
